---
title: "redo"
author: "Kourtney Burger"
date: "2025-01-15"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
    
---

# Load Packages

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE)

library(here)
library(tidyverse)
library(ggplot2)
library(sf)
library(marmap)
library(rnaturalearth)
library(rnaturalearthdata)
library(gganimate)
library(swfscDAS)
```

# Clean and Prep Data

## DASBRs

Note: There are too many DASBR points to plot these efficiently so I am
trying to to concatenate the data into hourly gps points (taking the
first point from each hour) 1. Load driftwatch functions and GPS
database

1.  Load driftwatch functions and GPS database

    ```{r}
    # Load driftwatch functions here - note: line 39 of this code is commented out, only works if full driftwatch directory is saved to this repo
    source(here('code/driftwatch/DriftWatchFunctions.R'))

    # Load GPS database, may need to download updated version from driftwatch google drive if new drifts aren't included
    db <- here('data/GPS/DASBRs/SPOTGPS_Logger.sqlite3')
    ```

2.  Select what data you want to plot

    ```{r}
    #All CalCurCEAS data
    DASBR_gps <- getDbDeployment(db)
    DASBR_gps <- filter(DASBR_gps, grepl('CalCurCEAS_', DriftName))
    ```

    For testing, subset the data

    ```{r}
    # DASBR_gps <- getDbDeployment(db, drift = c('CalCurCEAS_006', 'CalCurCEAS_007', 'CalCurCEAS_008', 'CalCurCEAS_009', 'CalCurCEAS_010', 'CalCurCEAS_011', 'CalCurCEAS_012', 'CalCurCEAS_013', 'CalCurCEAS_014', 'CalCurCEAS_015'))
    ```

3.  Average GPS data - 1 point per drift per hour

    ```{r}
    #Subset to only include DriftName, UTC, latitude, and longitude
    DASBR_gps <- DASBR_gps %>% 
      select(UTC,Latitude,Longitude,DriftName)

    DASBR_hourly_gps <- DASBR_gps %>%
      # Convert UTC to POSIXct format
      mutate(UTC = as.POSIXct(UTC, format = "%Y-%m-%d %H:%M:%S")) %>% 
      # 
      mutate(UTC = floor_date(UTC, "hour")) %>% 
      # Group by drift and the hour from UTC
      group_by(DriftName, UTC) %>% 
      # Average the data so there is one point per hour per drift
      summarize(
        Latitude = mean(Latitude, na.rm = TRUE), 
        Longitude = mean(Longitude, na.rm = TRUE), 
      )

    # Change DriftName to general name to match gliders and ship
    DASBR_hourly_gps <- DASBR_hourly_gps %>% rename(
      ID = DriftName
    )
    ```

## Gliders

1.  Load and combine glider data

    ```{r}
    # Load 3 glider gps files
    Glider_sg639 <- read.csv(here('data/GPS/Gliders/sg639_CalCurCEAS_Sep24_simpleSurfaceGPS.csv'))
    Glider_sg679 <- read.csv(here('data/GPS/Gliders/sg679_CalCurCEAS_Sep24_simpleSurfaceGPS.csv'))
    Glider_sg680 <- read.csv(here('data/GPS/Gliders/sg680_CalCurCEAS_Aug24_simpleSurfaceGPS.csv'))

    # Add ID column to each dataframe
    Glider_sg639 <- Glider_sg639 %>% 
      mutate(ID = 'sg639')

    Glider_sg679 <- Glider_sg679 %>% 
      mutate(ID = 'sg679')

    Glider_sg680 <- Glider_sg680 %>% 
      mutate(ID = 'sg680')

    # Combine into one dataframe
    glider_GPS <- bind_rows(Glider_sg639, Glider_sg679, Glider_sg680)
    ```

2.  Clean glider data - we do not need to average these because there
    are fewer points

    ```{r}
    # Only interested in GPS points at start of dive, remove end of dive points
    glider_GPS <- glider_GPS %>% 
      filter(label == "start")

    # correct datetime format
    glider_GPS$dateTime_UTC <- as.POSIXct(glider_GPS$dateTime_UTC, format = "%d-%b-%Y %H:%M:%S", tz = "UTC")

    # Subset to only include glider ID, datetime, latitude, and longitude
    glider_GPS <- glider_GPS %>% 
      select(ID, dateTime_UTC,latitude,longitude)

    # change column names
    glider_GPS <- glider_GPS %>% rename(
      ID = ID,
      UTC = dateTime_UTC,
      Latitude = latitude,
      Longitude = longitude
    )
    ```

## Ship Track

Selene prepped ship track data from DAS files (see [Selene's
code](https://github.com/NMFS-PAM-Glider/glider-CalCurCEAS/blob/main/R/shipEffortTrack.R))

```{r}
ship <- read.csv(here("data/GPS/Ship Track/ship_effort.csv"))

# make sure datetime format is correct
ship$DateTime <- as.POSIXct(ship$DateTime, format = "%Y-%m-%d %H:%M:%S", tz = "UTC")

# Subset to only include an ID column, datetime, latitude, and longitude
shipTrack_GPS <- ship %>% 
  select(line_num, DateTime, Lat, Lon)

# add ID column with Ship for all
shipTrack_GPS <- shipTrack_GPS %>% 
  mutate(ID = "Ship")

# change column names
shipTrack_GPS <- shipTrack_GPS %>% 
  rename(ID = ID,
         UTC = DateTime,
         Latitude = Lat,
         Longitude = Lon,
         LineNumber = line_num
         )
# remove line number for consistency purposes
# shipTrack_GPS <- shipTrack_GPS %>%
#   mutate(LineNumber = NULL)

# remove NA rows (2 rows had no time or latitude)
shipTrack_GPS <- na.omit(shipTrack_GPS)
```

## Combine DASBRs & Gliders

```{r}
# Excluding ship track so I can control that layer independently on map
AllGPS <- bind_rows(DASBR_hourly_gps, glider_GPS)
```

# Plot Data

## Prep Map Data/Layers

```{r}
# Convert DASBR and glider tracklines to sf object and set a standardized coordinate reference system
GPS_sf <- st_as_sf(AllGPS, coords = c("Longitude","Latitude"), crs = 4326)

# Convert GPS points to linestrings (NOTE: this excludes time as a variable)
GPS_sf_lines <- GPS_sf %>%
  arrange(ID) %>%
  group_by(ID) %>%
  summarise(geometry = st_sfc(st_linestring(st_coordinates(geometry)[, 1:2]))) %>%
  ungroup()

# Convert ship trackline to sf object and set a standardized coordinate reference system
shipTrack_GPS_sf <- st_as_sf(shipTrack_GPS, coords = c("Longitude", "Latitude"), crs = 4326)


# Extract the max and min coordinates of the data to set plot limits (NOTE: not tacking ship track into account for bounds)
bbox <- st_bbox(GPS_sf)
xlim_vals <- c(bbox["xmin"]-1, bbox["xmax"]+1)
ylim_vals <- c(bbox["ymin"], bbox["ymax"])


# Load bathymetric data for the plot limits selected above, +- 1 unit of lat and lon to make map boundaries extend
bathy <- getNOAA.bathy(lon1 = bbox["xmin"]-1, lon2 = bbox["xmax"]+1, 
                       lat1 = bbox["ymin"]-1, lat2 = bbox["ymax"]+1, 
                       resolution = 1)

# Convert bathymetry data object to a dataframe
bathy_df <- fortify.bathy(bathy)

# Load coastline and land data and convert it to an sf object
coastline_sf <- ne_download(scale = 10, type = 'coastline', category = 'physical', returnclass = "sf")
land <- ne_countries(scale = "medium", returnclass = "sf")
# cities ~ borrowed from Taiki's driftwatch code
POI <- tibble::tribble(
    ~Name, ~Latitude, ~Longitude,
#    "San Diego", 32.71068967391705, -117.17147162885448,
#    "Santa Barbara",34.407205041229595, -119.69269808900013,
#    'Ventura',34.250263795418434, -119.26720606042934,
#   'Morro Bay',35.36841835524968, -120.86325292077103,
#    'Monterey Bay',36.604218252060306, -121.89240128825472,
#    'Santa Cruz',36.96225980624226, -122.00212520928149,
#    'Half Moon Bay',37.50293801397416, -122.48765584637566,
    'San Francisco',37.813095434735914, -122.50037485469521,
#    'Bodega Bay',38.30982199529412, -123.05611099859385,
#    'Fort Bragg',39.42849849826603, -123.81309923266699,
#    'Shelter Cove',40.02403071506129, -124.06607534915634,
    'Eureka',40.806299478998056, -124.1807826182425,
#    'Crescent City',41.74692081633374, -124.19223894744171,
    'Point Conception',34.4483491615287, -120.47193766943991,
#    'Point Arena',38.91093890217707, -123.71170879559632,
#    'Cape Mendocino', 40.438268949326925, -124.40971460611878
#    'Coos Bay', 43.371729, -124.216278,
    'Newport', 44.63205, -124.044233,
    'Astoria',  46.181591, -123.830038,
#    'Wesport', 46.885821, -124.100979, 
#    'La Push', 47.902640, -124.602445
)
POI_sf <- st_as_sf(POI, coords = c("Longitude", "Latitude"), crs = 4326)


# Making sure all the mapping data is set to the same coordinate reference system (WGS84) (redundant for gps data but just making sure)
GPS_sf_lines <- st_set_crs(GPS_sf_lines, 4326) # GPS data
coastline_sf <- st_set_crs(coastline_sf, 4326) # coastline
land_sf <- st_set_crs(land, 4326) # land


# Prep points for adding labels to map. Each drift will be labeled on the first point
label_points <- GPS_sf %>%
  group_by(ID) %>%
  slice(1) %>%
  ungroup()

label_points <- label_points %>% # Extract coordinates for labeling
  mutate(
    x = st_coordinates(.)[,1],  # Extract the longitude (X)
    y = st_coordinates(.)[,2]   # Extract the latitude (Y)
  )


# Add new column with shortened ID for mapping
GPS_sf <- GPS_sf %>%
  mutate(ID_short = sub("CalCurCEAS_(\\d+)", "D\\1", ID), ID_short = sub("^D0+", "D", ID_short))

GPS_sf_lines <- GPS_sf_lines %>% 
  mutate(ID_short = sub("CalCurCEAS_(\\d+)", "D\\1", ID), ID_short = sub("^D0+", "D", ID_short))

label_points <- label_points %>% 
  mutate(ID_short = sub("CalCurCEAS_(\\d+)", "D\\1", ID), ID_short = sub("^D0+", "D", ID_short))
```

## Static Basemap

### Grouped by ID

```{r}
# Plot grouped by ID (i.e. dasbr, glider, ship)
p_ID <- ggplot() +
  # Add map layer
  geom_raster(data = bathy_df, aes(x = x, y = y, fill = z)) + # bathy data
  scale_fill_gradient(low = "#333333", high = "white") + # change bathy to grey scale
  geom_sf(data = land_sf) + # land layer
  geom_sf(data = coastline_sf, color = "black", size = 0.5) + # coastline layer
  geom_sf(data = POI_sf) + # cities
  geom_text(data = POI, aes(x = Longitude, y = Latitude, label = Name), size = 2, color = "#333333", nudge_x = 1) + # add city labels
  geom_sf(data = shipTrack_GPS_sf, aes(alpha = 0.3, linetype = "dashed"), size = 0.5) + # Ship track
  geom_sf(data = GPS_sf, aes(color = as.factor(ID), group = ID, linetype = "dashed"), size = 0.5) + # DASBRs and Gliders
  coord_sf(xlim = xlim_vals, ylim = ylim_vals) + # correct axis limits to match coordinate bounding box created above
  
  # Fix plot labels 
  labs(x = "Longitude", y = "Latitude", title = "DASBRs, gliders, & ship tracks", color = "ID") +
  
  # Tidy plot
  guides(
  # fill = guide_legend(title = "Depth (m)"), # adds a depth legend, include and comment out line below if you want this 
  fill = guide_none(), # removes depth legend, comment out line above if you want this
  color = guide_legend(title = "Drift ID"), # changes legend title
  linetype = "none", # removes the linetype legend 
  alpha = "none") +  
  theme(legend.position = "right",
        legend.margin = margin(t = 200), 
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10)) +
#  scale_color_viridis_d() + # change color scale 
  theme_minimal() # simplify plot

# save simple static plot
ggsave(here('figs/DASBR_Glider_Ship_SimpleStaticplot.png'), bg = "white")
```

### Grouped by Time

```{r}
# Plot grouped by time
p_UTC <- ggplot() +
  # Add map layer
  geom_raster(data = bathy_df, aes(x = x, y = y, fill = z)) + # bathy data
  scale_fill_gradient(low = "#333333", high = "white") + # change bathy to grey scale
  geom_sf(data = land_sf) + # land layer
  geom_sf(data = coastline_sf, color = "black", size = 0.5) + # coastline layer
  geom_sf(data = POI_sf) + # cities
  geom_text(data = POI, aes(x = Longitude, y = Latitude, label = Name), size = 2, color = "#333333", nudge_x = 1) + # add city labels
  geom_sf(data = shipTrack_GPS_sf, aes(color = UTC, alpha = 0.3, linetype = "dashed", size = 0.1), size = 0.1) + # Ship track
  geom_sf(data = GPS_sf, aes(color = UTC, group = ID, linetype = "dashed"), size = 0.5) + # DASBRs and Gliders
  coord_sf(xlim = xlim_vals, ylim = ylim_vals) + # correct axis limits to match coordinate bounding box created above
  geom_text(data = label_points, aes(x = x, y = y, label = ID_short), size = 2, color = "black", check_overlap = TRUE) +  # label first point
  
  # Fix plot labels 
  labs(x = "Longitude", y = "Latitude", title = "DASBRs, gliders, & ship tracks") +
  
  # Tidy plot
  guides(
  # fill = guide_legend(title = "Depth (m)"), # adds a depth legend, include and comment out line below if you want this 
  fill = guide_none(), # removes depth legend, comment out line above if you want this
  color = guide_colorbar(title = "Date (UTC)", barheight = unit(8, "cm"), ticks = TRUE), 
  color = guide_legend(title = "Date (UTC)"), # changes legend title
  linetype = "none", # removes the linetype legend 
  alpha = "none") +  
  scale_color_datetime(labels = scales::date_format("%Y-%m-%d"), breaks = scales::date_breaks("10 days")) +  # Show breaks every 10 days
  theme(legend.position = "right",
        legend.margin = margin(t = 200), 
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10)) +
  theme_minimal() # simplify plot

# save simple static plot
ggsave(here('figs/DASBR_Glider_Ship_SimpleStaticplot_UTC.png'), bg = "white")
```

## Animated Map

```{r}
# Animated plot by date
p_UTC + 
  transition_time(UTC) +
  labs(title = "Day: {frame_time}") +
  shadow_mark(alpha = 0.3, size = 0.5)

# Save animation
# anim_save(here('figs/300f_ship_tracks_animation.gif'), nframes = 100, width = 384,  height = 1080)
```

# Oregon zoomed in version

## Prep Data

```{r}
# DASBRs
DASBR_gps <- getDbDeployment(db, drift = c('CalCurCEAS_006', 'CalCurCEAS_007', 'CalCurCEAS_008', 'CalCurCEAS_009', 'CalCurCEAS_010', 'CalCurCEAS_011', 'CalCurCEAS_012', 'CalCurCEAS_013', 'CalCurCEAS_014', 'CalCurCEAS_015'))
#Subset to only include DriftName, UTC, latitude, and longitude
DASBR_gps <- DASBR_gps %>% 
  select(UTC,Latitude,Longitude,DriftName)

DASBR_hourly_gps <- DASBR_gps %>%
  # Convert UTC to POSIXct format
  mutate(UTC = as.POSIXct(UTC, format = "%Y-%m-%d %H:%M:%S")) %>% 
  # 
  mutate(UTC = floor_date(UTC, "hour")) %>% 
  # Group by drift and the hour from UTC
  group_by(DriftName, UTC) %>% 
  # Average the data so there is one point per hour per drift
  summarize(
    Latitude = mean(Latitude, na.rm = TRUE), 
    Longitude = mean(Longitude, na.rm = TRUE), 
  )

# Change DriftName to general name to match gliders and ship
DASBR_hourly_gps <- DASBR_hourly_gps %>% rename(
  ID = DriftName
)



# Gliders loaded above


# Excluding ship track so I can control that layer independently on map
AllGPS <- bind_rows(DASBR_hourly_gps, glider_GPS)
GPS_sf <- st_as_sf(AllGPS, coords = c("Longitude","Latitude"), crs = 4326)

# BBox
# Extract the max and min coordinates of the data to set plot limits (NOTE: not tacking ship track into account for bounds)
bbox <- st_bbox(GPS_sf)
xlim_vals <- c(bbox["xmin"]-1, bbox["xmax"]+1)
ylim_vals <- c(bbox["ymin"], bbox["ymax"])


# Subset ship track based on coordinate limits
shipTrack_GPS <- shipTrack_GPS %>%
  filter(Longitude >= xlim_vals[1] & Longitude <= xlim_vals[2] & 
         Latitude >= ylim_vals[1] & Latitude <= ylim_vals[2]) 
shipTrack_GPS_sf <- st_as_sf(shipTrack_GPS, coords = c("Longitude", "Latitude"), crs = 4326)
```

## Plot Data

### By ID

```{r}
ggplot() +
  # Add map layer
  geom_raster(data = bathy_df, aes(x = x, y = y, fill = z)) + # bathy data
  scale_fill_gradient(low = "#333333", high = "white") + # change bathy to grey scale
  geom_sf(data = land_sf) + # land layer
  geom_sf(data = coastline_sf, color = "black", size = 0.5) + # coastline layer
  geom_sf(data = POI_sf) + # cities
  geom_text(data = POI, aes(x = Longitude, y = Latitude, label = Name), size = 2, color = "#333333", nudge_x = .5) + # add city labels
  geom_sf(data = shipTrack_GPS_sf, aes(alpha = 0.3, linetype = "dashed"), size = 0.5) + # Ship track
  geom_sf(data = GPS_sf, aes(color = as.factor(ID), group = ID, linetype = "dashed"), size = 0.5) + # DASBRs and Gliders
  coord_sf(xlim = xlim_vals, ylim = ylim_vals) + # correct axis limits to match coordinate bounding box created above
  
  # Fix plot labels 
  labs(x = "Longitude", y = "Latitude", title = "DASBRs, gliders, & ship tracks", color = "ID") +
  
  # Tidy plot
  guides(
  # fill = guide_legend(title = "Depth (m)"), # adds a depth legend, include and comment out line below if you want this 
  fill = guide_none(), # removes depth legend, comment out line above if you want this
  color = guide_legend(title = "Drift ID"), # changes legend title
  linetype = "none", # removes the linetype legend 
  alpha = "none") +  
  theme(legend.position = "right",
        legend.margin = margin(t = 200), 
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10)) +
#  scale_color_viridis_d() + # change color scale 
  theme_minimal() # simplify plot

# save simple static plot
ggsave(here('figs/Oregon_DASBR_Glider_Ship_Staticplot.png'), bg = "white")
```

### By Time

```{r}
ggplot() +
  # Add map layer
  geom_raster(data = bathy_df, aes(x = x, y = y, fill = z)) + # bathy data
  scale_fill_gradient(low = "#333333", high = "white") + # change bathy to grey scale
  geom_sf(data = land_sf) + # land layer
  geom_sf(data = coastline_sf, color = "black", size = 0.5) + # coastline layer
  geom_sf(data = POI_sf) + # cities
  geom_text(data = POI, aes(x = Longitude, y = Latitude, label = Name), size = 2, color = "#333333", nudge_x = .5) + # add city labels
  geom_sf(data = shipTrack_GPS_sf, aes(color = UTC, alpha = 0.3, linetype = "dashed", size = 0.1), size = 0.1) + # Ship track
  geom_sf(data = GPS_sf, aes(color = UTC, group = ID, linetype = "dashed"), size = 0.5) + # DASBRs and Gliders
  coord_sf(xlim = xlim_vals, ylim = ylim_vals) + # correct axis limits to match coordinate bounding box created above
  geom_text(data = label_points, aes(x = x, y = y, label = ID_short), size = 2, color = "black", check_overlap = TRUE) +  # label first point
  
  # Fix plot labels 
  labs(x = "Longitude", y = "Latitude", title = "DASBRs, gliders, & ship tracks") +
  
  # Tidy plot
  guides(
  # fill = guide_legend(title = "Depth (m)"), # adds a depth legend, include and comment out line below if you want this 
  fill = guide_none(), # removes depth legend, comment out line above if you want this
  color = guide_colorbar(title = "Date (UTC)", barheight = unit(8, "cm"), ticks = TRUE), 
  color = guide_legend(title = "Date (UTC)"), # changes legend title
  linetype = "none", # removes the linetype legend 
  alpha = "none") +  
  scale_color_datetime(labels = scales::date_format("%Y-%m-%d"), breaks = scales::date_breaks("10 days")) +  # Show breaks every 10 days
  theme(legend.position = "right",
        legend.margin = margin(t = 200), 
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10)) +
  theme_minimal() # simplify plot

# save simple static plot
ggsave(here('figs/Oregon_DASBR_Glider_Ship_Staticplot_UTC.png'), bg = "white")
```
